---
title: "p8105_hw5_ez2384"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r Load Packages & Set Seed, message=FALSE}

  library(tidyverse)
  library(dplyr)
  library(tidyr)
  library(ggplot2)

```

### Problem 1
  /

### Problem 2
```{r Fix Parameters}
  
  n = 30
  sigma = 5

```

```{r Create Simulator Function}

  sim_func = function(sample_size, population_mean, population_std) {
    sim_data = tibble(sample = rnorm(n = sample_size, mean = population_mean, sd = population_std))
    stats  = sim_data %>% summarize(mu_hat = mean(sample))
    t_test = broom::tidy(t.test(sim_data$sample, mu=0)) %>% rename(mu_hat = estimate) %>% select(mu_hat, p.value)
    return(t_test)
  }

```

```{r Perform Simulation, cache=TRUE, message=FALSE, warning=FALSE}

  simulation_results =  tibble(mu = 0:6) %>% mutate(data = map(.x=mu, ~ rerun(5000, sim_func(sample_size=n, population_mean=.x, population_std=sigma)) %>% bind_rows()))

```

```{r Power Results & Plot showing the mu-power relationship, message=FALSE, warning=FALSE}

  power_results = simulation_results %>% mutate(power = map_dbl(.x=data, ~ mean(.x$p.value < 0.05))) %>% select(mu, power)
  
  ggplot(power_results, aes(x = mu, y = power)) +
    geom_line(size=0.3) +
    geom_point(size=0.5) +
    labs(
      title = "Power of the Test vs. True Value of μ",
      x = "True Value of μ",
      y = "Power (Proportion of Null Rejections)"
    ) +
    theme_light() +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks = 0:6) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.05))

```

#### Discussion on plot of true population mean against statistical power   
- When μ=0, the power is close to 0.05, which matches the significance level (alpha) of 0.05. This is expected because when the null hypothesis is true, the probability of rejecting it (Type I error rate) is equal to the significance level.    
- As μ moves further away from 0, the power increases steadily. This is because as the true mean μ diverges from the null hypothesis mean of 0, the effect size increases, making it easier for the test to detect a significant difference. Consequently, the likelihood of rejecting the null hypothesis rises.    
- Asymptotic Power (Power Approaching 1): At higher values of μ (around 4 to 6), the power approaches 1. This indicates that the test is almost always able to detect the difference from the null hypothesis when the true mean is substantially different from 0.   
- Conclusion: This plot correctly reflects the association between effect size (distance between the true μ and the null hypothesis mean) and power as it shows that 
Small effect sizes result in low power whereas larger effect sizes result in higher power. This result confirms that the test is more powerful as the effect size increases,

```{r}

  average_mu_results =  simulation_results %>% 
                        mutate( avg_mu_hat_all=map_dbl(.x=data, ~ mean(.x$mu_hat)), avg_mu_hat_rejected=map_dbl(data, ~ mean(.x$mu_hat[.x$p.value < 0.05])) ) %>%
                        select( mu, avg_mu_hat_all, avg_mu_hat_rejected )
  
  ggplot(average_mu_results, aes(x = mu)) +
    geom_line(aes(y = avg_mu_hat_all, color = "All Samples"), size = 0.35) +
    geom_point(aes(y = avg_mu_hat_all, color = "All Samples"), size = 0.5) +
    geom_line(aes(y = avg_mu_hat_rejected, color = "Rejected Samples"), size = 0.35) +
    geom_point(aes(y = avg_mu_hat_rejected, color = "Rejected Samples"), size = 0.5) +
    labs(
      title = "Average Estimate of μ̂ vs True Value of μ",
      x = "True Value of μ",
      y = "Average Estimate of μ̂",
      color = "Sample Type"
    ) +
    theme_light() +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_manual(values = c("All Samples" = "blue", "Rejected Samples" = "red")) +
    scale_x_continuous(breaks = 0:6) +
    scale_y_continuous(breaks = 0:6)

```

#### Discussion on plot of average mu_hat against true mu   
- No, the sample average μ_hat across tests for which the null hypothesis is rejected is not approximately equal to the true value of μ, especially when μ is close to the null hypothesis value (0).
- When μ is close to 0, only samples with sample means μ_hat that are far from 0 will produce a p-value below 0.05, leading to rejection of the null hypothesis. These extreme values inflate the average of μ_hat among the rejected samples.   
- By focusing only on rejected tests, we are selecting samples that deviate more from the null hypothesis mean (0), resulting in a higher average μ_hat than if we included all samples. This selection bias skews the average upwards for the "Rejected Samples" group.

### Problem 3   
```{r Loading Dataset}

  prob3 = read.csv("./homicide-data.csv") %>% mutate( city_state=paste(city, state, sep=", ") )
  total_homicides_per_city=prob3 %>% group_by(city) %>% summarize(total_homicides=n(), unsolved_homicides=sum(disposition == "Closed without arrest" | disposition == "Open/No arrest"))

```

#### Description of Raw Data    
- The data included information on a total of 52179 criminal homicides over the past decade in 50 of the largest cities within U.S. including the killing location, whether arrested, demographic details of each victim.   

```{r Function for Prob3, message=FALSE, warning=FALSE}

  baltimore_data = total_homicides_per_city %>% filter(city=="Baltimore")
  prop.test(baltimore_data$unsolved_homicides, baltimore_data$total_homicides) %>% broom::tidy() %>% 
                                                                                   mutate(confidence_interval=paste("(", conf.low, ", ",conf.high, ")")) %>% 
                                                                                   rename(point_estimate=estimate) %>% 
                                                                                   select(point_estimate, confidence_interval)
  prob3_func <- function(data) {
    data %>%
      mutate(prop_test = map2(unsolved_homicides, total_homicides, ~ prop.test(.x, .y) %>% broom::tidy())) %>%
      unnest(prop_test) %>%
      select(city, proportion_unsolved = estimate, lower_bound = conf.low, upper_bound = conf.high) %>% 
      mutate(confidence_interval = paste("(", lower_bound, ", ", upper_bound, ")"))
  }
  
  city_results = prob3_func(total_homicides_per_city)
  city_results %>% select(-lower_bound, -upper_bound)

```

```{r, message=FALSE, warning=FALSE}

  city_results <- city_results %>%
    arrange(desc(proportion_unsolved)) %>%
    mutate(city = factor(city, levels = city))
  
  ggplot(city_results, aes(x = city, y = proportion_unsolved)) +
    geom_point(size = 0.35, color = "blue") +
    geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.35, color = "red") +
    labs(
      title = "Proportion of Unsolved Homicides by City",
      x = "City",
      y = "Proportion of Unsolved Homicides"
    ) +
    theme_light() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
      axis.text.y = element_text(size = 10)
    )

```

